% thesis?

\documentclass[12pt]{report}
\usepackage{baththesis}

\usepackage{amsmath, amssymb, amsfonts, url, natbib, bm, rotating, multirow, graphicx, psfrag, appendix}

 
% Shortcuts
% Probability
\newcommand{\prob}[1]{\mathbb{P}\left[ #1 \right]}
% fprime
\newcommand{\fprime}{f^\prime(z)}
% phi inverse
\newcommand{\phiinv}{\varphi^{-1}}
\newcommand{\varphiinv}{\varphi^{-1}}
% use other phi by default
%\renewcommand{\phi}{\varphi}
%transpose
\newcommand{\tr}[1]{#1^{\text{T}}}
% diagonal
\newcommand{\diag}{\text{diag}}
% call \times \cross
\newcommand{\cross}{\times}
% shortcut for ij^{th}
\newcommand{\ijth}{$ij^{\text{th}}$}

% lazy citation correction...
\newcommand{\citeb}{\citet}
\renewcommand{\cite}{\citealp}

% for corrections, get it to put p. infront
\newcommand{\pref}{p. \pageref}

% references
% figure reference command
\newcommand{\fig}[1]{figure \ref{#1}}
% Figure reference command
\newcommand{\Fig}[1]{Figure \ref{#1}}
% table reference command
\newcommand{\tabref}[1]{table \ref{#1}}
% Table reference command
\newcommand{\Tabref}[1]{Table \ref{#1}}
% equation reference command
\newcommand{\eqn}[1]{(\ref{#1})}
% section references
\newcommand{\secref}[1]{section \ref{#1}}
\newcommand{\Secref}[1]{Section \ref{#1}}

\newcommand{\appref}[1]{appendix \ref{#1}}
\newcommand{\Appref}[1]{Appendix \ref{#1}}


% LaTeX laziness!
\newcommand{\bc}{\begin{center}}
\newcommand{\ec}{\end{center}}
\newcommand{\bn}{\begin{enumerate}}
\newcommand{\en}{\end{enumerate}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{eqnarray}}
\newcommand{\ee}{\end{eqnarray}}
\newcommand{\bes}{\begin{eqnarray*}}
\newcommand{\ees}{\end{eqnarray*}}
\newcommand{\expect}[1]{\mathbb{E}\left[ #1 \right]}

% words
\newcommand{\sch}{Schwarz-Christoffel}
\newcommand{\tprs}{thin plate regression spline}
\newcommand{\tprss}{thin plate regression splines}

% whatever I call MDS+TP approach
\newcommand{\mdsap}{MDS+RS}
% whatever I call the MDS with duchon splines approach
\newcommand{\mdsds}{MDS+DS}
% whatever I call the R package for this
\newcommand{\mdspack}{\texttt{msg}}

\title{On smooth models for complex domains and distances}
\author{David Lawrence Miller}
\degree{Doctor of Philosophy}
\department{Department of Mathematical Sciences}
\degreemonthyear{September 2011}

\norestrictions

\begin{document}

\maketitle

%\begin{quotation}
%``The major justification for the existence of any form of advanced voluntary organised education should be that it enables users to distort time and location in the learning process.'' 
%
%-- Cedric Price
%\end{quotation}



\begin{abstract} 
Spline smoothing is a popular technique for creating maps of a spatial phenomenon. Most smoothers use the Euclidean metric to measure the distance between data. This approach is flawed: phenomena do not conform to Euclidean geometry in their spatial distribution. Both natural and man-made barriers partition space and models should take this into account. The first part of this thesis develops a finite area smoother which does not suffer from this problem when the shape of the area is complex. It begins by rejecting the use of the \sch\ transform as a method for morphing complex domains due to its squashing of space. From there a method based on preserving within-area distances using multidimensional scaling is developed. High dimensional projections of the data are necessary to avoid a loss of ordering in the points. To smooth reliably in high dimensions Duchon splines are used. The model developed rivals the current best finite area method in prediction error terms and fits easily into larger models. Finally, the utility of projection methods to smooth general distances is explored.

The second part of the thesis concerns distance sampling, a widely used set of methods for estimating the abundance of biological populations. The work presented here introduces mixture formulation for the detection function used to model the probability of detection. The use of mixture models leads to flexible but monotonic detection functions, avoiding the unrealistic shapes which conventional methods are prone to. These new models are then applied to several existing, problematic data sets.

\newpage

\begin{center}
\section*{Acknowledgements}
\end{center}
I would first like to thank my supervisor Simon Wood for three years of guidance and assistance in both technical and presentational matters. At many points my own pessimism would have lead to me abandoning an approach; the work presented here is testament to his continued input.  My thanks also go to Simon Shaw who helped improve the work greatly via careful readings and comments. I would also like to thank the Engineering and Physical Sciences Research Council (EPSRC) for their financial support.

Without the work of my collaborators Len Thomas, Giampiero Marra and Luca Zanin a large portion of the thesis would not exist. In particular, I am indebted to my former officemate Giampiero Marra for a great many useful discussions over the past three years. In addition, Mark Taylor provided extremely useful comments on a first draft of the work in chapter \ref{chap-it}.

Going further back in time, to before the work in this thesis had begun, I must thank the Centre for Research into Environmental and Ecological Modelling (CREEM) at the University of St Andrews (and Steve Buckland specifically) for employing me in summer placements for the four years that I was an undergraduate. Without the encouragement and inspiration of those working at CREEM, I am certain that I would not have embarked on a PhD.

On a personal note, Elle Dodd's continued moral support and encouragement (as well as patience and tolerance!) through the highs and lows of the past three years have been essential for both of our sanities. Finally I would like to thank my parents, without whom this would have been (quite literally) impossible.
\end{abstract}



\tableofcontents

\listoffigures

\listoftables

\part{Finite area smoothing}

\chapter{Introduction to finite area smoothing}

\input{intro/intro}

\chapter{Modelling the spatiotemporal distribution of the incidence of resident foreign population in Italy}

\input{it/sp}

\chapter{Using the \sch\ transform to morph domains for finite area smoothing}
%\begin{quotation}
%``Failure is \textit{always} an option.'' 
%
%-- Adam Savage
%\end{quotation}

\input{sc/sc}

\chapter{Multidimensional scaling for domain-dependent finite area smoothing}

\input{mds/mds}

\chapter{Generalized distance smoothing}
%\begin{quotation}
%The three virtues of a programmer:
%\begin{enumerate}
%\item Laziness - The quality that makes you go to great effort to reduce overall energy expenditure. It makes you write labor-saving programs that other people will find useful, and document what you wrote so you don't have to answer so many questions about it. Hence, the first great virtue of a programmer. Also hence, this book. See also impatience and hubris.
%\item Impatience - The anger you feel when the computer is being lazy. This makes you write programs that don't just react to your needs, but actually anticipate them. Or at least pretend to. Hence, the second great virtue of a programmer. See also laziness and hubris.
%\item Hubris - Excessive pride, the sort of thing Zeus zaps you for. Also the quality that makes you write (and maintain) programs that other people won't want to say bad things about. Hence, the third great virtue of a programmer. See also laziness and impatience.
%\end{enumerate}
%-- Larry Wall, Randal L. Schwartz and Tom Christiansen. 
%\end{quotation}
%\begin{quotation}
%``The three virtues of a programmer: laziness, impatience and hubris.''
%
%-- Larry Wall, Randal L. Schwartz and Tom Christiansen. 
%\end{quotation}

\input{gds/gds}

\chapter{Conclusion}
\input{fasend/fasend}

\part{Distance sampling}

\chapter{Introduction to distance sampling}
\input{intro/intro-ds}


\chapter{Mixture model distance sampling detection functions}
% "it's another laser in your intergalactic battle cruiser" - Len Thomas

\input{mix/mix}

% start the appendices
\appendix

%\chapter{Properties of the within-area distance algorithm}
%
%\input{app-WAD/app-WAD}
%

\chapter{Derivatives of the likelihood for mixture model detection functions}

\input{app-mixderivs/mixderivs}


\bibliographystyle{chicago}
\bibliography{bib-hell}

\chapter*{Corrections}
\input{corrections.tex}

\end{document}
